### 2018-05-14 source/Ch2-intro/calibration.tex
reorganised source files

\section{Calibration Methods in Radio Interferometry}\label{section.calibration}
\pg
In this section, we will discuss the implementation of interferometric array calibration. Our analysis is based on the RIME formalism, described in \cref{section.RIME}. One key metric of calibration quality is the \emph{dynamic range} (DR). High dynamic ranges mean that a high contrast has been obtained, and fainter sources can be reached. Dynamic range is defined as the following:
\begin{equation}\label{eq.DR}
DR = \frac{flux_{max}}{\max(\sigma_{thermal},\sigma_{artefacts})}
\end{equation}
where $flux_{max}$ is the flux of the brightest source, $\sigma_{thermal}$ the thermal noise in the image, and $\sigma_{artefacts}$ the noise associated with calibration artefacts.

\pg
The distinction between these two noise sources is crucial; one can never go `below noise' for a given observation, no matter the quality of calibration. Astronomers typically observe for longer periods of time in order to reduce $\sigma_{thermal}$ in their images, but this will not reduce the artefacts caused by poor calibration solutions. However, uncorrected Direction-Dependent Effects will not go away on their own, no matter how long the integration time.

\pg
There are three `generations' of calibration methods, of increasing complexity. I shall describe them in terms of the RIME, showing how each generation increases in generality to account for more exotic effects. For the sake of clarity, I shall henceforth refer to them interchangeably as `nth-generation calibration' or `nGC' methods.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing a calibrator before and after each observation run to find a gain factor and offset error\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{here}.}

\pg
Phase calibration in the 1GC era `proper' was not done, as engineers were capable of ensuring phase stability in contemporary interferometers. Phase was thus calculated relative to a fixed frame of reference, usually the central antenna of a 3-antenna array.

@@

\pg
Second-generation calibration methods (2GC methods) are defined by their use of \emph{self-calibration}\footnote{On the discovery of self-calibration and its evolution in parallel to adaptive optics, I highly recommend the chapter titled "The Almost Serendipitous Discovery of Self-Calibration'' in "\href{http://library.nrao.edu/public/collection/02000000000280.pdf}{\citep{serendipitous}}.}, commonly referred to as \emph{self-cal} (\citepads{1984ARA&A..22...97P}). As described in \cref{section.RIME.FullSky.CVZ}, this method can only be deployed if the brightness matrix of the sky is the same for all baselines\footnote{This is not to say that the brightness matrix must contain only point sources, but rather that moving our interferometer 500m to the East should not change its measured visibilities} (i.e. that we are not affected by direction-dependent effects).

\pg
The seeds of self-calibration (and adaptive optics) is a paper published in the era of 1GC by Jennison (\citepads{1958MNRAS.118..276J}, expanding on his PhD work, which was published in 1951). With sufficient signal-to-noise, he showed that \emph{phase closure} could be calculated and errors due to the atmosphere thus mitigated.

@@

\pg
[ TODO: put adaptive optics diagram here maybe]

@@

\pg
By extending this idea to VLBI\footnote{e.g.  \citepads{1974ApJ...193..293R}}, \emph{amplitude closure} was introduced to the field along with phase closure \citepads[as described in][]{1983Sci...219...51R}. Astronomers were quick to apply these methods to interferometers in general.

@@

\pg
In this section, I will discuss the approach taken by [cite DDF and kMS paper when it comes out] and [cite prefactor paper if/when it is out], \emph{facet-based calibration}. This method has an imaging equivalent \citepads[see][and associated papers]{2017arXiv171202078T}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

\pg
The crucial insight of this approach consists

>>>>

\section{Calibration Methods in Radio Interferometry}\label{section.calibration}
\pg
In this section, we will discuss the implementation of interferometric array calibration. Our analysis is based on the RIME formalism, described in \cref{section.RIME}. One key metric of calibration quality is the \emph{dynamic range}, mentioned in Sec. \ref{section.clean}. High dynamic ranges mean that a high contrast has been obtained, and fainter sources can be reached. Here, we redefine dynamic range as follows
\begin{equation}\label{eq.DR}
DR = \frac{flux_{max}}{\max(\sigma_{thermal},\sigma_{artefacts})}
\end{equation}
where $flux_{max}$ is the flux of the brightest source, $\sigma_{thermal}$ the thermal noise in the image, and $\sigma_{artefacts}$ the noise associated with calibration artefacts. This gives a definition for dynamic range useful even in fields with a single source. We can thus use it as a metric for calibration quality specifically.

\pg
The distinction between these two noise sources is crucial; one can never go `below noise' for a given observation, no matter the quality of calibration. Astronomers typically observe for longer periods of time in order to reduce $\sigma_{thermal}$ in their images, but this will not reduce the artefacts caused by poor calibration solutions. Uncorrected Direction-Dependent Effects will not go away on their own, no matter how long the integration time. Similarly, there is a limit to how much improving calibration will improve the final image: eventually, more data is required to drive noise down.

\pg
There are three `generations' of calibration methods, of increasing complexity. We will describe them in terms of the RIME, showing how each generation increases in generality to account for more exotic effects. They are referred to interchangeably as `nth-generation calibration' or `nGC' methods.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing a calibrator before and after each observation run to find a gain factor and offset error\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{the open-loop calibration techniques page of analog.com}.}

\pg
Phase calibration in the 1GC era `proper' was not necessary, as engineers were capable of ensuring adequate phase stability in contemporary interferometers. Phase was thus calculated relative to a fixed frame of reference, usually the central antenna of a 3-antenna array.

@@

\pg
Second-generation calibration methods (2GC methods) are defined by \emph{self-calibration}\footnote{On the discovery of self-calibration and its evolution in parallel to adaptive optics, see the chapter titled "The Almost Serendipitous Discovery of Self-Calibration'' in "\href{http://library.nrao.edu/public/collection/02000000000280.pdf}{\citep{serendipitous}}.}, commonly referred to as \emph{self-cal} (\citepads{1984ARA&A..22...97P}). As described in \cref{section.RIME.FullSky.CVZ}, this method can only be deployed if the brightness matrix of the sky is the same for all baselines\footnote{This is not to say that the brightness matrix must contain only point sources, but rather that moving our interferometer 500m to the East should not change its measured visibilities} (i.e. that we are not affected by direction-dependent effects).

\pg
The first instance of self-calibration (and adaptive optics in radio interferometry) is a paper published in the era of 1GC by Jennison \citepads{1958MNRAS.118..276J}, expanding on his PhD work, which was published in 1951). With sufficient signal-to-noise, he showed that \emph{phase closure} could be calculated and errors due to the atmosphere thus mitigated.

@@

@@

\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/selfcal.png}}
\caption{\label{fig.selfcal} Representation of self-calibration as adaptive optics}
\end{figure}

\pg
By extending this idea to VLBI\footnote{e.g.  \citepads{1974ApJ...193..293R}}, \emph{amplitude closure} was introduced to the field along with phase closure \citepads[as described in][]{1983Sci...219...51R}. These quantities are immune to antenna-based effects.%Astronomers were quick to apply these methods to interferometers in general.

@@

\pg
In this section, I will discuss the approach taken by \textcolor{red}{[cite DDF and kMS paper when it comes out]} and \textcolor{red}{[cite prefactor paper if/when it is out]}, \emph{facet-based calibration}. This method has an imaging equivalent \citepads[see][and associated papers]{2017arXiv171202078T}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. % The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

\begin{figure}[ht]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/bootes-image.png}}
\caption{\label{fig.facets} Voronoi facets, as implemented in DDFacet. Image is made before application of direction-dependent correction. This is an image of the Bootes field, made by Cyril Tasse.}
\end{figure}

\pg
When performing first and second-generation calibration, one solves for a set of Jones matrices or parameters thereof. Direction-independent calibration then consists of applying these calibration solutions to the visibilities directly. However, while these solutions will be correct in the direction of the calibrator source, they are increasingly likely to be wrong as a function of distance from said calibrator source. The difference between the true gains in a given direction and the gains estimated from the calibrator source are called \emph{differential gains}.

\pg
Why then not solve for a set of calibration solutions for each source in the calibration model, iteratively adding new sources to this model as they become above the noise as calibration artefacts - introduced by poor calibration - decrease? There are two main reasons for this: firstly, the signal-to-noise when calibrating using very faint source models will be poor. This means that gains will likely be extremely noisy, and applying them will result in increased calibration artefacts in the final image. Secondly, even by optimising for signal-to-noise and only using a few of the brightest sources in the field, the problem quickly becomes intractably large in terms of computing power required, provided that standard complex differentiation is used \citepads[see][]{2016ApJS..223....2V}. Wirtinger differentiation \citepads[see][and associated papers]{2014arXiv1410.8706T} does alleviate this somewhat, but is not implemented in all standard direction-dependent calibration software at the time of writing.

\pg
The crucial insight of a faceting approach is then to split the image into a set of \emph{facets}, solving for a set of calibration solutions for each of these facets. These solutions are then applied to the visibilities when mapping them onto the dirty image. The faceting in DDF is shown in \cref{fig.facets}. The facet distribution shown in \cref{fig.facets} encourages an even flux distribution in multiple facets, and ensures that no facet is so small that too little signal would be available within. As we can see from \cref{fig.DDcal.effect}, direction-dependent calibration can drastically improve the quality of wide-field images. It currently represents the state of the art in interferometric calibration.

\begin{figure}[h!]
\centering
\begin{subfigure}{\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/lofar-nocorr.png}}
\caption{\label{fig.lofar.noDDcal} Image of the Boötes field, made without direction-dependent calibration. Image by Cyril Tasse.}
\end{subfigure}
\begin{subfigure}{\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/lofar-corr.png}}
\caption{\label{fig.lofar.withDDcal} Image of the same Boötes field, made with direction-dependent calibration. Image by Cyril Tasse.}
\end{subfigure}
%\hfill
\caption{\label{fig.DDcal.effect} Difference between imaging made with (Fig. \ref{fig.lofar.withDDcal})and without (Fig. \ref{fig.lofar.noDDcal}) third-generation calibration.}
\end{figure}

### 2018-05-14 source/Ch2-intro/imaging.tex
reorganised source files

\section{A Brief Introduction to Radio Astronomy}
\pg
Radio astronomy consists of observing the electromagnetic emission of astrophysical sources at very long wavelengths by means of radio antennas. These antennas measure a voltage proportional to variations in the electromagnetic field in all the directions they are sensitive to. Since astrophysical signals are weak, a good sensitivity is necessary. Achieving good sensitivity with radio antennas means having a very large collecting area - in this respect, they behave exactly the same way as optical telescopes. Similarly, when well-designed, they are diffraction-limited - this means that, once again like optical telescopes, their resolution is limited by their diameter.

@@

\pg
Finally, one type of weighting scheme of note is simple data flagging. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a unit weight. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

>>>>

\section{A Brief Introduction to Radio Astronomy}
\pg
Radio astronomy consists of observing the electromagnetic emission of astrophysical\footnote{And, famously, attempting to observe ``intelligent" sources \citepads[cf]{2017AAS...22911604E}} sources at very long wavelengths by means of radio antennas. These antennas measure a voltage proportional to variations in the electromagnetic field in all the directions they are sensitive to. Since astrophysical signals are weak, a good sensitivity is necessary. Achieving good sensitivity with radio antennas means having a very large collecting area - in this respect, they behave exactly the same way as optical telescopes. Similarly, when well-designed, they are diffraction-limited - this means that, once again like optical telescopes, their resolution is limited by their diameter.

@@

\pg
Finally, one type of weighting scheme of note is simple data flagging \citepads[cf.]{2010ascl.soft10017O}. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a unit weight. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

### 2018-05-14 source/Ch2-intro/science.tex
reorganised source files

\subsection{Free-Free Radiation}
\pg
Free-free or ``bremsstrahlung" (German for ``braking") radiation occurs when the trajectory of a high-energy charged particle is deflected by an electric field. This non-thermal emission mechanism is the dominant mechanism in HII regions (which contain ionised hydrogen), where star formation has previously taken place. It is called free-free emission because it is produced by free electrons scattering off ionised hydrogen without being captured. % as shown in Fig. \ref{plot.freefree}
%\begin{figure*}[!h]
%\centering
%\includegraphics[width=0.3\textwidth]{images/bremsstrahlung.png}
%\caption{\label{plot.freefree} Schematic showing the physical process leading to free-free emission. As we can see, the electron is scattered, but not captured. Source: \url{https://thephysicsbehind.com/2015/04/16/x-ray-tubes/}}
%\end{figure*}

>>>>

\subsection{Free-Free Radiation}
\pg
Free-free or ``bremsstrahlung" (``braking") radiation occurs when the trajectory of a high-energy charged particle is deflected by an electric field. This non-thermal emission mechanism is the dominant mechanism in HII regions (which contain ionised hydrogen), where star formation has previously taken place. It is called free-free emission because it is produced by free electrons scattering off ionised hydrogen without being captured. % as shown in Fig. \ref{plot.freefree}
%\begin{figure*}[!h]
%\centering
%\includegraphics[width=0.3\textwidth]{images/bremsstrahlung.png}
%\caption{\label{plot.freefree} Schematic showing the physical process leading to free-free emission. As we can see, the electron is scattered, but not captured. Source: \url{https://thephysicsbehind.com/2015/04/16/x-ray-tubes/}}
%\end{figure*}

### 2018-05-14 source/Ch3-RIME/RIME.tex
reorganised source files

\pg
We shall use the conventional coordinate system and notations (\citetads{2011A&A...527A.106S}, \citetads{2001isra.book.....T}, etc). We thus have a $z$ axis pointing towards the phase centre, and an antenna $p$ at coordinates $\U_p=(u_p,v_p,w_p)$. The phase difference between $p$ and phase centre (where, by definition, $\U=0$) for a signal arriving from direction $\pmb{\sigma}$ is then:
\begin{equation}
k_p=2\pi i (( u_p l + v_p m + w_p (n-1) ))
\end{equation}

@@

\pg 
Having done this, we can now introduce a scalar \emph{K-Jones} matrix to represent the effect of this phase delay. Since it is scalar, it will commute with all other Jones matrices (see \citetads{2011A&A...527A.106S}). It can be be written as:
\begin{equation}
\Kjones_p = e^{-i k_p}  = e^{-2\pi i ( u_p l + v_p m + w_p (n-1) )}
\end{equation}

@@

\pg
$\Jones(t,\nu)$ describes the propagation of an electromagnetic signal through various physical processes. This signal has a complex phase, which is variable in both time and frequency. Since it is a rotating vector, integrating over any time or frequency band will always result in a \emph{net loss} in amplitude in the measured $\langle\Vis\rangle$. This mechanism is commonly referred to as \emph{time/bandwidth width decorrelation} or \emph{smearing} (\cite{2011A&A...527A.106S}).

@@

\pg Smearing increases with baseline length ($\uvec_{pq}$) and distance from phase centre ($l,m$). The noise amplitude, however, does not decrease. This means that smearing results in a decrease in sensitivity, which becomes significant when attempting to do wide-field, high-resolution images (see section \textcolor{red}{[PLUG GROTH WORK SHAMELESSLY]}).

>>>>

\pg
We shall use the conventional coordinate system and notations \citepads[][etc]{2011A&A...527A.106S,2001isra.book.....T}. We thus have a $z$ axis pointing towards the phase centre, and an antenna $p$ at coordinates $\U_p=(u_p,v_p,w_p)$. The phase difference between $p$ and phase centre (where, by definition, $\U=0$) for a signal arriving from direction $\pmb{\sigma}$ is then:
\begin{equation}
k_p=2\pi i (( u_p l + v_p m + w_p (n-1) ))
\end{equation}

@@

\pg 
Having done this, we can now introduce a scalar \emph{K-Jones} matrix to represent the effect of this phase delay. Since it is scalar, it will commute with all other Jones matrices \citepads[see]{2011A&A...527A.106S}. It can be be written as:
\begin{equation}
\Kjones_p = e^{-i k_p}  = e^{-2\pi i ( u_p l + v_p m + w_p (n-1) )}
\end{equation}

@@

\pg
$\Jones(t,\nu)$ describes the propagation of an electromagnetic signal through various physical processes. This signal has a complex phase, which is variable in both time and frequency. Since it is a rotating vector, integrating over any time or frequency band will always result in a \emph{net loss} in amplitude in the measured $\langle\Vis\rangle$. This mechanism is commonly referred to as \emph{time/bandwidth width decorrelation} or \emph{smearing} \citepads{2011A&A...527A.106S}.

@@

\pg Smearing increases with baseline length ($\uvec_{pq}$) and distance from phase centre ($l,m$). The noise amplitude, however, does not decrease. This means that smearing results in a decrease in sensitivity, which becomes significant when attempting to do wide-field, high-resolution images (see \cref{section.EGS.hires}).

### 2018-05-14 source/Ch5-Strategy/strategy.tex
reorganised source files

\pg
Why, then, use this approach? One reason is, in fact, for the challenge itself. A high-resolution (matching HST resolution), wide-area (multiple degrees wide), high-quality radio survey has \emph{never} been achieved at the time of writing. Many surveys of the radio sky exist, of course, but they tend to combine only two of those three qualities. FIRST\citepads[see][and references therein]{2015ApJ...801...26H}, for example, has achieved tremendous results with a $4''$ resolution, relatively low sensitivity, and no short baselines (thereby missing diffuse emission and extended structure). The VLASS \citepads[see][and references therein]{2013arXiv1312.4602H} will aim to go to a $2.5''$ resolution, medium sensitivity, and few short baselines. %Finally, there will of course be the SKA1-MID survey once the SKA is operational; this survey is expected to provide excellent resolution, sensitivity, and good $uv$-coverage, making it sensitive to diffuse emission as well as point sources.

@@

\section{Imaging the Full Primary Beam}
\pg
Perform DID calibration at low resolutions (no international baselines) to get complete, approximatemodel of entire EGS field

\section{Test Decorrelation}
\pg
see effect of decorrelation on LOBOS sources around EGS as function of distance from 3c295 - talk about two sources of decorrelation (direction-dependent PSF, which is modelled, and gains changing with direction, which is unmodelled and will have an impact in image). See maximum impact of gain-decorrelation: we want flat decorrelation as function of distance from 3c295.

>>>>

\pg
Why, then, use this approach? One reason is, in fact, for the challenge itself. A high-resolution (matching HST resolution), wide-area (multiple degrees wide), high-quality radio survey has \emph{never} been achieved at the time of writing. Many surveys of the radio sky exist, of course, but they tend to combine only two of those three qualities. FIRST\citepads[see][and references therein]{2015ApJ...801...26H}, for example, has achieved tremendous results with a $4''$ resolution, relatively low sensitivity, and no short baselines (thereby missing diffuse emission and extended structure). The VLASS \citepads[see][and references therein]{2013arXiv1312.4602H} will aim to go to a $2.5''$ resolution, medium sensitivity, and few short baselines. There exist many LOFAR sky surveys \citepads[cf.]{2016MNRAS.460.2385W,2017A&A...598A.104S}, but they do not yet make use of international LOFAR, and are thus limited in resolution ($\sim 5''$).%Finally, there will of course be the SKA1-MID survey once the SKA is operational; this survey is expected to provide excellent resolution, sensitivity, and good $uv$-coverage, making it sensitive to diffuse emission as well as point sources.

@@

\pg
The first step towards acquiring a deep, high-resolution image of the Extended Groth Strip therefore consists of creating a good, high-resolution model of 3C295 at LOFAR frequencies. Creating this model is one of the key results of this thesis. Starting from a high-resolution VLA model of 3C295, subbands selected across the LOFAR bandwidth are self-calibrated in iteratively greater numbers until a given noise threshold is reached. The flux scale is then boot-strapped based on the results of \citet{arse}. This ensures that 3C295 will be adequately subtracted from all corrected visibilities - including international baselines - and thus that images made using these visibilities will not be polluted by sidelobes from 3C295.

\pg
The work done towards this goal can be found in \cref{section.3c295}, and its result - a high-resolution spectral model of 3C295 - is shown below.

\pg
\textcolor{red}{show key result here}

\section{Imaging the Full Primary Beam}
\pg
After creating a high-resolution model of 3C295 to calibrate the international stations, we must create a low-resolution model of all the sources in the primary beam to both improve final calibration quality and subtract sources outside of the Extended Groth Strip when creating the final high-resolution images.

\pg
Initial direction-independent calibration is done using the same pipeline as for the Lofar Two-metre Sky Survey \citepads[LOTSS,][]{2017A&A...598A.104S}. The killMS-DDF facet-based pipeline \citepads[cf]{2016ApJS..223....2V,2015MNRAS.449.2668S,2017arXiv171202078T} is then used to perform direction-dependent calibration for the entire LOFAR primary beam. This allows us to reach a noise threshold of \textcolor{red}{give noise threshold}.

\begin{figure}[h!]
\centering
\includegraphics[width=0.95\textwidth]{images/{lofar.widefield}.png}
\caption{\label{plot.EGS.lofar.widefield} Full direction-dependent calibrated image of the LOFAR primary beam. Image made without international stations.}
\end{figure}

\pg
This work, and its results, are shown in \cref{section.EGS.lowres}. Note that the AEGIS observations summarised in \cref{table.EGS.observation} do not cover the full LOFAR primary beam. The sources observed at low resolution are therefore matched with SDSS images. Overlays are shown to highlight the wealth of interesting sources, and their optical counterparts, which can be picked up as a byproduct of producing high-resolution maps of the radio sky. We select 12 particularly interesting sources and show the results of source matching for these candidates. Overlays of the same SDSS images with NVSS postage stamps are also shown for comparison. Many of the chosen sources do not lie within the Extended Groth Strip, and are therefore only imaged at low resolution, but follow-up images and observations could be made should the need arise.

\section{Test Decorrelation}
\pg
Before launching the EGS imaging run using all calibrated datasets, we measure the impact of direction-dependent effects on known sources in the field. Two direction-dependent effects are expected to dominate: decorrelation and differential gains. The first is modeled by the imager used, whereas the second is a function of the gain variability as a function of distance from the calibrator. Seven calibrator sources, lying at various distances and directions from the calibrator source, are examined. They are summarised in \cref{table.LOBOS.sources}, which is reproduced below.

\begin{table}[h!]
\begin{tabular}{ccccc}
\# & RA [hms]    & Dec [dms]   & Dist. from EGS [deg.] & Dist. from 3C295 [deg.] \\\hline
1  & 14:30:18.72 & 52:17:29.80 & 2.041                         & 2.904 \\
2  & 14:19:44.44 & 54:23:04.58 & 1.928                         & 2.517 \\ 
3  & 14:21:20.05 & 53:03:46.00 & 0.864                         & 1.743 \\
4  & 14:21:09.41 & 51:22:32.46 & 1.294                         & 1.728 \\
5  & 14:11:50.32 & 52:49:02.66 & 0.844                         & 0.619 \\
6  & 14:11:20.23 & 52:12:04.30 & 0.915                         & 0.000 \\
7  & 14:08:07.00 & 52:55:11.36 & 1.409                         & 0.869 \\
8  & 14:08:09.76 & 52:44:46.56 & 1.354                         & 0.680 \\
\end{tabular}
\caption{\label{table.LOBOS.sources1}Table recapitulating the positions of all 8 chosen calibrator sources, along with their distance from the observation phase centre (EGS phase centre) and calibrator phase centre (3C295), respectively.}
\end{table}
%\begin{figure}[h!]
%\includegraphics[width=0.8\linewidth]{images/EGS_LOBOS_scatterplot}
%\caption{Position of the calibrator sources around the EGS. Note that this is not projected properly onto the celestial sphere.}
%\label{bootes-coverage-image}
%\end{figure}

\pg
If directional gains are measured to have little or no impact, then direction-dependent calibration of the international stations will not be required. This is the ideal scenario, as it does not require the use of fringe fitting \citepads{1984ARA&A..22...97P}. If they are measured to have a significant impact, then direction-dependent calibration will be required for international LOFAR. We expect the angular scale of differential gain evolution to be of the order of a degree, and so most of the EGS should be relatively unaffected.

\pg
\textcolor{red}{show key result + some discussion here}

### 2018-05-14 source/Ch6-3C295/3c295.tex
reorganised source files

\chapter{Imaging 3C295 with International LOFAR}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}
\pg
Our aim in this section is to create a high-resolution model of 3C295, something that - as
of yet - does not exist. More specifically, we seek to create a model that will allow us to find
good phase-calibration solutions for LOFAR international baselines. While we also solve for
amplitude gains, we know that we will need to correct the total source flux in each frequency
subband as our initial spectral model is not necessarily correct.
We select 6 sub-bands out of the total LOFAR HBA bandwidth, evenly spread throughout the bandwidth as shown in Fig. \ref{fig.freqsamp}.

@@

\subsection{Data \& Observation Properties}
\pg
The dataset used for this PhD project is part of the LOFAR Surveys KSP Tier-1 survey, which consists of a number of 8-hour pointing covering as much of the sky visibile to LOFAR as possible. We analyse one of these pointings, an observation performed on the 28th of August 2014 and centred on the Extended Groth Strip. As part of the Tier-1 survey, it is an 8-hour-long observation. We limit ourselves to analysing only the HBA observation, meaning that we use 365 sub-bands which sample a total bandwidth of [give bandwidth]. We use all Core and Remote LOFAR stations, as well as some of the International stations which were online at the time (specifically the German stations 1-5 and 7, along with the Swedish and the British station).

>>>>

\chapter{Imaging 3C295 with International LOFAR}\label{section.3c295}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}
\pg
Our aim in this section is to create a high-resolution model of 3C295, something that - as
of yet - does not exist. More specifically, we seek to create a model that will allow us to find
good phase-calibration solutions for LOFAR international baselines. While we also solve for
amplitude gains, we know that we will need to correct the total source flux in each frequency
subband as our initial spectral model is not necessarily correct.
We select 6 sub-bands out of the total LOFAR HBA bandwidth, evenly spread throughout the bandwidth as shown in Fig. \ref{fig.freqsamp}.

@@

\subsection{Data \& Observation Properties}
\pg
The dataset used for this PhD project is part of the LOFAR Surveys KSP Tier-1 survey, which consists of a number of 8-hour pointings covering as much of the sky visibile to LOFAR as possible. We analyse one of these pointings, an observation performed on the 28th of August 2014 and centred on the Extended Groth Strip. As part of the Tier-1 survey, it is an 8-hour-long observation. We limit ourselves to analysing only the HBA observation, meaning that we use 365 sub-bands which sample a total bandwidth of [give bandwidth]. We use all Core and Remote LOFAR stations, as well as some of the International stations which were online at the time (specifically the German stations 1-5 and 7, along with the Swedish and the British station).

### 2018-05-14 source/Ch7-loresEGS/lowres-EGS.tex
reorganised source files

\subsection{Data \& Observation Properties}
\pg
The data used to make this image are the same as used to make the high-resolution model of 3C295 in \cref{section.3c295}, but we do not use the visibilities of baselines including international stations in our calibration or imaging. They were acquired through the LOFAR Long-Term Archive service. We do not phase-shift our data when imaging.

@@

\pg
An initial sky model is generated using the VLSSr \citepads[VLA Low-frequency Sky Survey - redux, ]{2012RaSc...47.0K04L}, WENSS \citepads[WEsterbork Northern Sky Survey]{1997A&AS..124..259R} and the NVSS \citepads[NRAO VLA Sky Survey, ]{1998AJ....115.1693C}. This model serves as the basis for facet-based self-calibration, performed using the killMS-DDF surveys pipeline \citepads{2017A&A...598A.104S}.
\pg
The direction-dependent calibration was done using \textcolor{red}{[N]} facets, spread as shown in \textcolor{red}{fig. make figure}. Some facets converge faster or better than others. Calibration is done solving for full, complex Jones matrices, rather than a physics-based or diagonal-and-rotation matrix [CHECK THAT THIS IS TRUE!]. Imaging is done [GIVE WEIGHTING, ETC].

@@

>>>>

%\subsection{Data \& Observation Properties}
%\pg
%The data used to make this image are the same as used to make the high-resolution model of 3C295 in \cref{section.3c295}, but we do not use the visibilities of baselines including international stations in our calibration or imaging. They were acquired through the LOFAR Long-Term Archive service. We do not phase-shift our data when imaging.

@@

\pg
An initial sky model is generated using the VLSSr \citepads[VLA Low-frequency Sky Survey - redux, ]{2012RaSc...47.0K04L}, WENSS \citepads[WEsterbork Northern Sky Survey]{1997A&AS..124..259R} and the NVSS \citepads[NRAO VLA Sky Survey, ]{1998AJ....115.1693C}. This model serves as the basis for facet-based self-calibration, performed using the killMS-DDF surveys pipeline \citepads{2017A&A...598A.104S}.

\begin{figure}[h!]
\centering
\includegraphics[width=0.95\textwidth]{images/{EGS.widefield.regions}.png}
\caption{\label{fig.EGS.widefield.regions} Full direction-dependent calibrated image of the LOFAR primary beam. Image made without international stations.}
\end{figure}

\pg
The direction-dependent calibration was done using 121 facets, spread as shown in \cref{fig.EGS.widefield.regions}. Some facets converge faster or better than others. Calibration is done solving for full, complex Jones matrices, rather than a physics-based or diagonal-and-rotation matrix [CHECK THAT THIS IS TRUE!].
The final image was imaged with a very small $uv$-cut, keeping only baselines longer than 100m. This is because shorter baselines were not used during calibration, and so discarded for imaging. The image resolution is $1.5''$, and the image was created using both quality-based weighting schemes and Briggs weighting, with a Briggs factor of $-0.5$. This means that the imaging was skewed slightly in favour of uniform weighting (weighting visibilities by local $uv$-density) rather than natural weighting (same weight for all visibilities). Imaging and calibration were done modelling the LOFAR beam, so the image gives integrated flux values rather than apparent flux values. Deconvolution was performed using SSD-clean with an external mask.

@@

\pg
We will now show images, overlays and cross-matching analysis \& results for the 12 chosen sources in the primary beam.
